---
title: "Chloropleth_Maps"
---

**Informative Maps**

Let’s look at informative maps using state-level housing data from the openintro package. In this tab, I use median home value by state as my numeric variable. For my categorical variable, I group states into Higher Cost and Lower Cost housing states depending on whether their median home value is above or below the national median.

To better explore the data, each variable is displayed in two forms: a static choropleth map and an interactive Choropleth map. This allows us to compare patterns across states while also interacting with the maps for more detail.

```{r}
#| message: false
#| warning: false
#| include: false

library(tidyverse)
library(maps)
library(sf)
library(leaflet)
library(openintro)
library(viridis)
library(htmltools)
```

```{r}
#| echo: false
# map polygons for static plots
states_polygon <- as_tibble(map_data("state")) |>
  select(region, group, order, lat, long)

# sf map for interactive plots
states_sf <- 
  read_sf("https://rstudio.github.io/leaflet/json/us-states.geojson") |>
  select(name, geometry)

# create state housing data
housing_data <- county_complete |>
  group_by(state) |>
  summarise(
    median_home_value = median(median_val_owner_occupied_2010, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    key = str_to_lower(state),
    cost_group = if_else(
      median_home_value >
        median(median_home_value, na.rm = TRUE),
      "Higher Cost",
      "Lower Cost"
    )
  )

# merge for static map
states_static <- states_polygon |>
  left_join(housing_data, by = c("region" = "key"))

# merge for interactive map
states_interactive <- states_sf |>
  mutate(key = str_to_lower(name)) |>
  left_join(housing_data, by = "key") |>
  filter(!(name %in% c("Alaska","Hawaii","Puerto Rico")))
```

**Numeric Variable: Median Home Value** **Static Map**

```{r}
#| echo: false
#| classes: center-map
ggplot(states_static,
       aes(long, lat, group = group)) +
  geom_polygon(aes(fill = median_home_value),
               color = "white") +
  coord_map() +
  theme_void() +
  scale_fill_viridis(
    labels = scales::dollar
  ) +
  labs(
    title = "Median Home Value by State",
    fill = "Home Value",
    caption = "Source: openintro package"
  )
```

**Insight:** From the map, we can see that housing prices are not the same across the country. Some states have higher median home values, while others have lower ones. This suggests that where a state is located may play a role in housing costs.

**Interactive Map**

```{r}
#| echo: false
pal_num <- colorNumeric(
  "viridis",
  domain = states_interactive$median_home_value
)

labels <- lapply(
  paste0(
    states_interactive$name,
    ": $",
    round(states_interactive$median_home_value,0)
  ),
  HTML
)

leaflet(states_interactive) |>
  addTiles() |>
  setView(-96, 37.8, 4) |>
  addPolygons(
    fillColor = ~pal_num(median_home_value),
    weight = 1,
    color = "white",
    fillOpacity = 0.7,
    label = labels
  ) |>
  addLegend(
    pal = pal_num,
    values = ~median_home_value,
    title = "Median Home Value",
    position = "bottomright"
  )
```

------------------------------------------------------------------------

**Categorical Variable: Housing Cost Group** **Static Map**

```{r}
#| echo: false
#| classes: center-map
ggplot(states_static,
       aes(long, lat, group = group)) +
  geom_polygon(aes(fill = cost_group),
               color = "white") +
  coord_map() +
  theme_void() +
  scale_fill_manual(
    values = c("Higher Cost" = "darkred",
               "Lower Cost" = "steelblue")
  ) +
  labs(
    title = "Higher vs Lower Cost Housing States",
    fill = "Category",
    caption = "Source: openintro package"
  )
```

**Insight:**

From the map, we can see that higher-cost states are more concentrated in coastal areas, while lower-cost states are more common in the South and Midwest. This shows that location can have a strong connection to housing prices across the country.

**Interactive Map**

```{r}
#| echo: false
pal_cat <- colorFactor(
  c("steelblue", "darkred"),
  domain = states_interactive$cost_group
)

labels2 <- lapply(
  paste0(
    states_interactive$name,
    ": ",
    states_interactive$cost_group
  ),
  HTML
)

leaflet(states_interactive) |>
  addTiles() |>
  setView(-96, 37.8, 4) |>
  addPolygons(
    fillColor = ~pal_cat(cost_group),
    weight = 1,
    color = "white",
    fillOpacity = 0.7,
    label = labels2
  ) |>
  addLegend(
    pal = pal_cat,
    values = ~cost_group,
    title = "Housing Group",
    position = "bottomright"
  )
```
